---
title: "Examples of the PlaneGeometry package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{examples}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(PlaneGeometry)
```

# Exeter point

The *Exeter point* is defined as follows on Wikipedia. 

Let $ABC$ be any given triangle. Let the medians through the vertices $A$, $B$, 
$C$ meet the circumcircle of triangle $ABC$ at $A'$, $B'$ and $C'$ respectively. 
Let $DEF$ be the triangle formed by the tangents at $A$, $B$, and $C$ to the 
circumcircle of triangle $ABC$. (Let $D$ be the vertex opposite to the side 
formed by the tangent at the vertex $A$, let $E$ be the vertex opposite to the 
side formed by the tangent at the vertex $B$, and let $F$ be the vertex 
opposite to the side formed by the tangent at the vertex $C$.) 
The lines through $DA'$, $EB'$ and $FC'$ are concurrent. 
The point of concurrence is the Exeter point of triangle $ABC$.

Let's construct it with the `PlaneGeometry` package. We do not need to 
construct the triangle $DEF$: it is the *tangential triangle* of $ABC$, and 
is provided by the `tangentialTriangle` method of the R6 class `Triangle`.

```{r}
A <- c(0,2); B <- c(5,4); C <- c(5,-1)
t <- Triangle$new(A, B, C)
circumcircle <- t$circumcircle()
centroid <- t$centroid()
medianA <- Line$new(A, centroid)
medianB <- Line$new(B, centroid)
medianC <- Line$new(C, centroid)
Aprime <- intersectionCircleLine(circumcircle, medianA)[[2]]
Bprime <- intersectionCircleLine(circumcircle, medianB)[[2]]
Cprime <- intersectionCircleLine(circumcircle, medianC)[[1]]
DEF <- t$tangentialTriangle()
lineDAprime <- Line$new(DEF$A, Aprime)
lineEBprime <- Line$new(DEF$B, Bprime)
lineFCprime <- Line$new(DEF$C, Cprime)
( ExeterPoint <- intersectionLineLine(lineDAprime, lineEBprime) )
# check whether the Exeter point is also on (FC')
lineFCprime$includes(ExeterPoint)
```

Let's draw a figure now.

```{r Exeter, fig.width=5, fig.height=5}
opar <- par(mar = c(0,0,0,0))
plot(0, 0, type = "n", asp = 1, xlim = c(-2,9), ylim = c(-6,7),
     xlab = NA, ylab = NA, axes = FALSE)
draw(t, lwd = 2, col = "black")
draw(circumcircle, lwd = 2, border = "cyan")
draw(Triangle$new(Aprime,Bprime,Cprime), lwd = 2, col = "green")
draw(DEF, lwd = 2, col = "blue")
draw(Line$new(ExeterPoint, DEF$A, FALSE, FALSE), lwd = 2, col = "red")
draw(Line$new(ExeterPoint, DEF$B, FALSE, FALSE), lwd = 2, col = "red")
draw(Line$new(ExeterPoint, DEF$C, FALSE, FALSE), lwd = 2, col = "red")
points(rbind(ExeterPoint), pch = 19, col = "red")
par(opar)
```


# Circles tangent to three circles

Let $\mathcal{C_1}$, $\mathcal{C_2}$ and $\mathcal{C_3}$ be three circles with 
respective radii $r_1$, $r_2$ and $r_3$ such that $r_3 < r_1$ and $r_3 < r_2$. 
How to construct some circles simultaneously tangent to these three circles?

```{r}
C1 <- Circle$new(c(0,0), 2)
C2 <- Circle$new(c(5,5), 3)
C3 <- Circle$new(c(6,-2), 1)
# inversion swapping C1 and C3 with positive power
iota1 <- inversionSwappingTwoCircles(C1, C3, positive = TRUE)
# inversion swapping C2 and C3 with positive power
iota2 <- inversionSwappingTwoCircles(C2, C3, positive = TRUE)
# take an arbitrary point on C3
M <- C3$center + c(C3$radius,0)
# invert it with iota1 and iota2
M1 <- iota1$invert(M); M2 <- iota2$invert(M)
# take the circle C passing through M, M1, M2
C <- Triangle$new(M,M1,M2)$circumcircle()
# take the line passing through the two inversion poles
cl <- Line$new(iota1$pole, iota2$pole)
# take the radical axis of C and C3
L <- C$radicalAxis(C3)
# let H bet the intersection of these two lines
H <- intersectionLineLine(L, cl)
# take the circle Cp with diameter [HO3]
O3 <- C3$center
Cp <- Circle$new((H+O3)/2, sqrt(c(crossprod(H-O3)))/2)
# get the two intersection points T0 and T1 of C3 with Cp
T0_and_T1 <- intersectionCircleCircle(C3, Cp)
T0 <- T0_and_T1[[1L]]; T1 <- T0_and_T1[[2L]]
# invert T0 with respect to the two inversions
T0p <- iota1$invert(T0); T0pp <- iota2$invert(T0)
# the circle passing through T0 and its two images is a solution
Csolution0 <- Triangle$new(T0, T0p, T0pp)$circumcircle()
# invert T1 with respect to the two inversions
T1p <- iota1$invert(T1); T1pp <- iota2$invert(T1)
# the circle passing through T1 and its two images is another solution
Csolution1 <- Triangle$new(T1, T1p, T1pp)$circumcircle()
```

```{r tangentCircles, fig.width=5, fig.height=5}
opar <- par(mar = c(0,0,0,0))
plot(0, 0, type = "n", asp = 1, xlim = c(-4,9), ylim = c(-4,9),
     xlab = NA, ylab = NA, axes = FALSE)
draw(C1, col = "yellow", border = "red")
draw(C2, col = "yellow", border = "red")
draw(C3, col = "yellow", border = "red")
draw(Csolution0, lwd = 2, border = "blue")
draw(Csolution1, lwd = 2, border = "blue")
par(opar)
```


# Filling the lapping area of two circles

Let two circles intersecting at two points. 
How to fill the lapping area of the two circles?

```{r lappingArea, fig.width=5, fig.height=5}
O1 <- c(2,5); circ1 <- Circle$new(O1, 2)
O2 <- c(4,4); circ2 <- Circle$new(O2, 3)

opar <- par(mar = c(0,0,0,0))
plot(0, 0, type = "n", asp = 1, xlim = c(0,8), ylim = c(0,8))
draw(circ1, border = "purple", lwd = 2)
draw(circ2, border = "forestgreen", lwd = 2)

intersections <- intersectionCircleCircle(circ1, circ2)
A <- intersections[[1]]; B <- intersections[[2]]
points(rbind(A,B), pch = 19, col = c("red", "blue"))

theta1 <- Arg((A-O1)[1] + 1i*(A-O1)[2]) 
theta2 <- Arg((B-O1)[1] + 1i*(B-O1)[2]) 
path1 <- Arc$new(O1, circ1$radius, theta1, theta2)$path()

theta1 <- Arg((A-O2)[1] + 1i*(A-O2)[2]) 
theta2 <- Arg((B-O2)[1] + 1i*(B-O2)[2]) 
path2 <- Arc$new(O2, circ2$radius, theta2, theta1)$path()

polypath(rbind(path1,path2), col = "yellow")

par(opar)
```


# Hyperbolic tessellation 

In the help page of the `Circle` R6 class (`?Circle`), we show how to draw 
a hyperbolic triangle with the help of the method 
`orthogonalThroughTwoPointsOnCircle()`. Here we will use this method to 
draw a hyperbolic tessellation.

```{r}
tessellation <- function(depth, Thetas0, colors){
  stopifnot(
    depth >= 3, 
    is.numeric(Thetas0),
    length(Thetas0) == 3L,
    is.character(colors), 
    length(colors) >= depth
  )
  
  circ <- Circle$new(c(0,0), 3)
  
  arcs <- lapply(seq_along(Thetas0), function(i){
    ip1 <- ifelse(i == length(Thetas0), 1L, i+1L)
    circ$orthogonalThroughTwoPointsOnCircle(Thetas0[i], Thetas0[ip1], 
                                            arc = TRUE)
  })
  inversions <- lapply(arcs, function(arc){
    Inversion$new(arc$center, arc$radius^2)
  })
  
  Ms <- vector("list", depth)
  
  Ms[[1L]] <- lapply(Thetas0, function(theta) c(cos(theta), sin(theta)))
  
  Ms[[2L]] <- vector("list", 3L)
  for(i in 1L:3L){
    im1 <- ifelse(i == 1L, 3L, i-1L)
    M <- inversions[[i]]$invert(Ms[[1L]][[im1]])
    attr(M, "iota") <- i
    Ms[[2L]][[i]] <- M
  }
  
  for(d in 3L:depth){
    n1 <- length(Ms[[d-1L]])
    n2 <- 2L*n1
    Ms[[d]] <- vector("list", n2)
    k <- 0L
    while(k < n2){
      for(j in 1L:n1){
        M <- Ms[[d-1L]][[j]]
        for(i in 1L:3L){
          if(i != attr(M, "iota")){
            k <- k + 1L
            newM <- inversions[[i]]$invert(M)
            attr(newM, "iota") <- i
            Ms[[d]][[k]] <- newM
          }
        }
      }
    }
  }
  
  # plot ####
  opar <- par(mar = c(0,0,0,0), bg = "black")
  plot(0, 0, type = "n", asp = 1, xlim = c(-4,4), ylim = c(-4,4))
  draw(circ, border = "white")
  invisible(lapply(arcs, draw, col = colors[1L], lwd = 2))
  
  Thetas <- lapply(
    rapply(Ms, function(M){
      Arg(M[1] + 1i*M[2])
    }, how="replace"),
    unlist)
  
  for(d in 2L:depth){
    thetas <- sort(unlist(Thetas[1L:d]))
    for(i in 1L:length(thetas)){
      ip1 <- ifelse(i == length(thetas), 1L, i+1L)
      arc <- circ$orthogonalThroughTwoPointsOnCircle(thetas[i], thetas[ip1],
                                                     arc = TRUE)
      draw(arc, lwd = 2, col = colors[d])
    }
  }
  
  par(opar)
  
  invisible()
}
```


```{r tessellation, fig.width=5, fig.height=5}
tessellation(
  depth = 5L, 
  Thetas0 = c(0, 2, 3.8), 
  colors = viridisLite::viridis(5)
)
```

Here is a version which allows to fill the hyperbolic triangles:

```{r}
tessellation2 <- function(depth, Thetas0, colors){
  stopifnot(
    depth >= 3, 
    is.numeric(Thetas0),
    length(Thetas0) == 3L,
    is.character(colors), 
    length(colors)-1L >= depth
  )
  
  circ <- Circle$new(c(0,0), 3)
  
  arcs <- lapply(seq_along(Thetas0), function(i){
    ip1 <- ifelse(i == length(Thetas0), 1L, i+1L)
    circ$orthogonalThroughTwoPointsOnCircle(Thetas0[i], Thetas0[ip1], 
                                            arc = TRUE)
  })
  inversions <- lapply(arcs, function(arc){
    Inversion$new(arc$center, arc$radius^2)
  })
  
  Ms <- vector("list", depth)
  
  Ms[[1L]] <- lapply(Thetas0, function(theta) c(cos(theta), sin(theta)))
  
  Ms[[2L]] <- vector("list", 3L)
  for(i in 1L:3L){
    im1 <- ifelse(i == 1L, 3L, i-1L)
    M <- inversions[[i]]$invert(Ms[[1L]][[im1]])
    attr(M, "iota") <- i
    Ms[[2L]][[i]] <- M
  }
  
  for(d in 3L:depth){
    n1 <- length(Ms[[d-1L]])
    n2 <- 2L*n1
    Ms[[d]] <- vector("list", n2)
    k <- 0L
    while(k < n2){
      for(j in 1L:n1){
        M <- Ms[[d-1L]][[j]]
        for(i in 1L:3L){
          if(i != attr(M, "iota")){
            k <- k + 1L
            newM <- inversions[[i]]$invert(M)
            attr(newM, "iota") <- i
            Ms[[d]][[k]] <- newM
          }
        }
      }
    }
  }
  
  # plot ####
  opar <- par(mar = c(0,0,0,0), bg = "black")
  plot(0, 0, type = "n", asp = 1, xlim = c(-4,4), ylim = c(-4,4))

  path <- do.call(rbind, lapply(rev(arcs), function(arc) arc$path()))
  # x <- unlist(lapply(paths, "[[", "x"))
  # y <- unlist(lapply(paths, "[[", "y"))
  polypath(path, col = colors[1L])
  
  invisible(lapply(arcs, function(arc){
    path1 <- arc$path()
    B <- arc$startingPoint()
    A <- arc$endingPoint()
    alpha1 <- Arg(A[1L] + 1i*A[2L])
    alpha2 <- Arg(B[1L] + 1i*B[2L])
    path2 <- Arc$new(c(0,0), 3, alpha1, alpha2)$path()
    #x <- c(path1$x, path2$x); y <- c(path1$y, path2$y)
    polypath(rbind(path1,path2), col = colors[2L])
  }))
  
  Thetas <- lapply(
    rapply(Ms, function(M){
      Arg(M[1] + 1i*M[2])
    }, how="replace"),
    unlist)
  
  for(d in 2L:depth){
    thetas <- sort(unlist(Thetas[1L:d]))
    for(i in 1L:length(thetas)){
      ip1 <- ifelse(i == length(thetas), 1L, i+1L)
      arc <- circ$orthogonalThroughTwoPointsOnCircle(thetas[i], thetas[ip1],
                                                     arc = TRUE)
      path1 <- arc$path()
      B <- arc$startingPoint()
      A <- arc$endingPoint()
      alpha1 <- Arg(A[1L] + 1i*A[2L])
      alpha2 <- Arg(B[1L] + 1i*B[2L])
      path2 <- Arc$new(c(0,0), 3, alpha1, alpha2)$path()
      #x <- c(path1$x, path2$x); y <- c(path1$y, path2$y)
      polypath(rbind(path1,path2), col = colors[d+1L])
    }
  }

  draw(circ, border = "white")
  
  par(opar)
  
  invisible()
}
```


```{r tessellation2, fig.width=5, fig.height=5}
tessellation2(
  depth = 5L, 
  Thetas0 = c(0, 2, 3.8), 
  colors = viridisLite::viridis(6)
)
```


# Director circle of an ellipse 

Let's draw the [director circle](https://en.wikipedia.org/wiki/Director_circle) 
of an ellipse. We start by constructing the minimum bounding box of the ellipse.

```{r}
ell <- Ellipse$new(c(1,1), 5, 2, 30)
majorAxis <- ell$diameter(0)
minorAxis <- ell$diameter(pi/2)
v1 <- (majorAxis$B - majorAxis$A) / 2
v2 <- (minorAxis$B - minorAxis$A) / 2
# sides of the minimum bounding box
side1 <- majorAxis$translate(v2)
side2 <- majorAxis$translate(-v2)
side3 <- minorAxis$translate(v1)
side4 <- minorAxis$translate(-v1)
# take a vertex of the bounding box
A <- side1$A
# radius of the director circle
R <- sqrt(c(crossprod(A - ell$center)))
# director circle
circ <- Circle$new(ell$center, R)
```

Now let's take a tangent $T_1$ to the ellipse, construct the half-line 
directed by $T_1$ with origin the point of tangency, determine the intersection 
point of this half-line with the director circle, and draw the perpendicular 
$T_2$ of $T_1$ passing by this intersection point. 
Then $T_2$ is another tangent to the ellipse.

```{r, message=FALSE}
T1 <- ell$tangent(0.3)
halfT1 <- T1$clone(deep = TRUE)
halfT1$extendA <- FALSE
I <- intersectionCircleLine(circ, halfT1, strict = TRUE)
T2 <- T1$perpendicular(I)
```

```{r directorCircle, fig.width=5, fig.height=5}
opar <- par(mar=c(0,0,0,0))
plot(NULL, type = "n", asp = 1, 
     xlim = c(-3,6), ylim = c(-5,7), xlab = NA, ylab = NA)
# draw the ellipse
draw(ell, col = "blue")
# draw the bounding box
draw(side1, lwd = 2, col = "green")
draw(side2, lwd = 2, col = "green")
draw(side3, lwd = 2, col = "green")
draw(side4, lwd = 2, col = "green")
# draw the director circle
draw(circ, lwd = 2, border = "red")
# draw the two tangents
draw(T1); draw(T2)
# restore the graphical parameters
par(opar)
```


# Playing with Steiner chains

The `PlaneGeometry` package has a function `SteinerChain` which generates 
a Steiner chain of circles.

## Elliptical Steiner chain

By applying an affine transformation to a Steiner chain, we can get an 
elliptical Steiner chain.

```{r}
c0 <- Circle$new(c(3,0), 3) # exterior circle
circles <- SteinerChain(c0, 3, -0.2, 0.5)
# take an ellipse 
ell <- Ellipse$new(c(-4,0), 4, 2.5, 140)
# take the affine transformation which maps the exterior circle to this ellipse
f <- AffineMappingEllipse2Ellipse(c0, ell)
# take the images of the Steiner circles by this transformation
ellipses <- lapply(circles, f$transformEllipse)
```

```{r ellipticalSteiner, fig.width=6, fig.height=4}
opar <- par(mar = c(0,0,0,0))
plot(NULL, type = "n", asp = 1, xlim = c(-8,6), ylim = c(-4,4), 
     xlab = NA, ylab = NA, axes = FALSE)
# draw the Steiner chain
invisible(lapply(circles, draw, lwd = 2, col = "blue"))
draw(c0, lwd = 2)
# draw the elliptical Steiner chain
invisible(lapply(ellipses, draw, lwd = 2, col = "red", border = "forestgreen"))
draw(ell, lwd = 2, border = "forestgreen")
par(opar)
```

## Nested Steiner chains

We can choose the exterior circle of the Steiner chain. Therefore, given a 
circle of a Steiner chain, we can nest another Steiner chain in this circle.

```{r}
c0 <- Circle$new(c(3,0), 3) # exterior circle
circles <- SteinerChain(c0, 3, -0.2, 0.5)
# Steiner chain for each circle, except the small one (it is too small)
chains <- lapply(circles[1:3], function(c0){
  SteinerChain(c0, 3, -0.2, 0.5)
})
```

```{r nestedSteiner, fig.width=5, fig.height=5}
opar <- par(mar = c(0,0,0,0))
plot(NULL, type = "n", asp = 1, xlim = c(0,6), ylim = c(-4,4), 
     xlab = NA, ylab = NA, axes = FALSE)
# draw the big Steiner chain
invisible(lapply(circles, draw, lwd = 2, border = "blue"))
draw(c0, lwd = 2)
# draw the nested Steiner chain
invisible(lapply(chains, function(circles){
  lapply(circles, draw, lwd = 2, border = "red")
}))
par(opar)
```
